---
title: "Homework 3"
author: "Rose Killian"
output: github_document
---

### Loading libraries and data

```{r message= FALSE}
library(tidyverse)
library(ggridges)
library(viridis)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

library(p8105.datasets)

```

# Problem 1
```{r}
data("instacart")
```

The Instacart dataset contains nearly 1.4 million observations of items ordered from Instacart by individuals in New York City in 2017.

* How many aisles are there. and which aisles are the most items ordered from?

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))
```

There are 134 aisles.The top 3 aisles ordered from are fresh vegetables, fresh fruits, and packaged vegetables.


* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  filter(n_obs > 10000) %>% 
  ggplot(aes(x = reorder(aisle, -n_obs), y = n_obs)) +
  geom_col() + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(
    x = "Aisle",
    y = "Number of items ordered"
  )
```

Fresh vegetables and fresh fruits have the most items ordered by far with each category having almost twice the number of items ordered as the third highest, packaged vegetables and fruits, has. 

* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  summarize(
    product_max = max(product_name),
    n_obs = n()
  ) %>% 
  knitr::kable()
```

This is not what I want. How do I get top 3 per group????

* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  mutate(mean_hod = round(mean(order_hour_of_day))) %>%
  group_by(product_name) %>% 
  select(product_name, order_dow, mean_hod) %>% 
  distinct() %>%
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hod"
  ) %>% 
  relocate(product_name, `0`, `1`, `2`, `3`, `4`, `5`, `6`) %>% 
  rename("Sunday" = `0`, "Monday" = `1`, "Tuesday" = `2`, "Wednesday" = `3`, "Thursday" = `4`, "Friday" = `5`, "Saturday" = `6`) %>% 
  knitr::kable()
```

*needs comment*



# Problem 2

```{r}
data("brfss_smart2010")
```

First, some data cleaning:

```{r}
brfss = 
brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health",
         response %in% c("Poor", "Fair", "Good", "Very Good", "Excellent")) %>% 
  mutate(response = factor(response, c("Poor", "Fair", "Good", "Very Good", "Excellent")))
```

* In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss %>%
  filter(year == "2002") %>% 
  group_by(locationabbr) %>% 
  summarise(n_obs = n()) %>% 
  mutate(locations = n_obs / 4) %>% 
  filter(locations > 6)

brfss %>%
  filter(year == "2010") %>% 
  group_by(locationabbr) %>% 
  summarise(n_obs = n()) %>% 
  mutate(locations = n_obs / 4) %>% 
  filter(locations > 6)
```

In 2002, 6 states had 7 or more observation locations: CT, FL, MA, NC, NJ, and PA.

In 2010, 14 states had 7 or more observation locations: CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA.

* Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss %>% 
  select(year, locationabbr, response, data_value, geo_location) %>% 
  filter(response == "Excellent") %>% 
  group_by(locationabbr, year) %>% 
  mutate(data_avg = mean(data_value)) %>% 
  select(year, locationabbr, data_avg) %>% 
  ggplot(aes(x = year, y = data_avg, color = locationabbr)) +
  geom_line() +
  theme(legend.position = "right")
```

*needs comment*

* Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss %>% 
  filter(locationabbr == "NY", year %in% c("2006", "2010")) %>% 
  group_by(year, locationdesc) %>% 
  ggplot(aes(x = response, y = data_value, color = locationdesc)) +
  geom_point() +
  facet_grid(. ~ year)
```

The overall `data_value` range of each response type is fairly consistent across 2006 and 2010. The range of values for fair responses is wider in 2010 and the max `data_value` is highest in 2006 among the good responses.

# Problem 3

Import and tidy data:
```{r}
accel_df = 
  read_csv("data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    weekday = recode(day, "Sunday" = 0, "Saturday" = 0, "Monday" = 1, "Tuesday" = 1, "Wednesday" = 1, "Thursday" = 1, "Friday" = 1)) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity"
  )
```
The data set `accel_df` contains `r nrow(accel_df)` observations of accelerometer data from a 63 year-old male collected every minute over 5 weeks. The variable `week` denotes what week from 1 to 5 the observation is from, `day_id` denotes the day since start of the study, `day` indicates the day of the week, `weekday` indicates if the day was a weekday or not (1 for weekday, 0 for weekend), `minute` indicates the minute of the day the observation occurs, starting at midnight, and `activity` is the activity count for each observation.

* Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_df %>% 
  group_by(day_id) %>% 
  mutate(tot_activity = sum(activity)) %>% 
  select(day_id, day, tot_activity) %>% 
  distinct() %>% 
  knitr::kable()
```

I don't really see any trends here.

* Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel_df %>% 
  group_by(day_id) %>% 
  ggplot(aes(x = minute, y = activity, color = day)) +
  geom_point(alpha = .8) +
  theme(axis.text.x = element_text(angle = 90))
```

The x-axis is a mess and I do not know why.